Viscous Transport in Eroding Porous Media

### 2019-10-02 docs/transport/erosion_transport.tex
small changes in language
NM

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
First, consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\cite{mit-spa2017}
performed a similar three-dimensional experiment using a second-order
quadrature method.  They initialized the \st{eroding} body \st{so that}
with its center located $1.5$ radii above the solid wall.  In two
dimensions, the trapezoid rule is spectrally accurate, and an analogous
initial configuration is possible. However, unless the number of
discretization points is increased, the trapezoid rule becomes obsolete
if the eroding body is initialized closer to the solid wall. Therefore,
to simulate erosion near a solid wall at reasonable levels of
discretization, we use our hybrid method that combines the trapezoid
rule and the Barycentric quadrature rule.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies have
vanished.  Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity, and the
erosion rate is proportional to the vorticity.  Initially, several of
the eroding bodies are closer to the outer wall than the $5h$ threshold
required to guarantee that the trapezoid rule achieves machine
precision.  In particular, the distance between bodies 1, 6, 13, and 15
and the outer wall is $1.3h$, $2.9h$, $2.8h$, and $1.3h$, respectively,
where $h$ is the arclength spacing of the outer wall $\Gamma$.  In
addition, the distance between several pairs of eroding bodies,
including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too small to be
accurately resolved with the trapezoid rule.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved at the prescribed resolution, and erosion can
be simulated until all the bodies have vanished.

@@

Similar to our previous work, erosion causes the space between nearly
touching bodies to quickly expand, and flat faces develop along the
regions of near contact.  This qualitative behavior is present in
Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
others.  However, by resolving the interaction between bodies that are
much closer together than in our previous work, we observe that, at
least initially, very little erosion occurs between certain pairs of
bodies.  For instance the opening between bodies 1 \& 6, 3 \& 9, and 5
\& 13 grow much slower than the opening between bodies 15 \& 16.  A
common feature of the pores that grow slowly is that they are nearly
perpendicular to the main flow direction, and this results in a small
erosion rate.

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) is roughly linear which indicates
that the flow rate can be written as a power law.  Using a line of best
fit, the flow rate is approximately 
\begin{align} 
  U \approx \left(\frac{t}{t_f}\right)^{3.43},
\end{align}
which is the dashed line in Figure~\ref{fig:Eroding20flowrate}(b).

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 gaps, 258 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 gaps}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 gaps}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 gaps}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the pore
sizes of 100 eroding bodies at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840, 0.3194)$.}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A Single Body Close to a Wall}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
First, consider a single eroding body close to a solid wall with the
shear flow
\begin{align}
  \UU(\xx)=\left[
  \begin{array}{c}
    y+1 \\ 0
  \end{array}
  \right],
\end{align}
imposed on the outer wall. Mitchell and Spagnolie~\cite{mit-spa2017}
performed a similar three-dimensional experiment using a second-order
quadrature method.  They initialized the \nick{\st{eroding} body \st{so that}
with its center located} $1.5$ radii above the solid wall.  In two
dimensions, the trapezoid rule is spectrally accurate, and an analogous
initial configuration is possible. However, unless the number of
discretization points is increased, the trapezoid rule becomes obsolete
if the eroding body is initialized closer to the solid wall. Therefore,
to simulate erosion near a solid wall at reasonable levels of
discretization, we use our hybrid method that combines the trapezoid
rule and the Barycentric quadrature rule.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{20 Bodies at a Medium Porosity}
\label{sec:Eroding20}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We consider 20 eroding bodies that are discretized with $N_\iin=256$
points.  The background flow is the Hagen-Poiseuille flow
\begin{align}
  \UU(\xx)=U \left[
  \begin{array}{c}
    1-y^2 \\ 0
  \end{array}
  \right],
\end{align}
where the flow rate $U$ is chosen so that the pressure drop from $x=-2$
to $x=2$ is held fixed at 8. Therefore, $U=1$ once all the bodies have
vanished.  Snapshots of the bodies at four equispaced times are shown in
Figure~\ref{fig:Eroding20vort}.  The color is the vorticity, \nick{which is proportional to the erosion rate on solid boundaries}.
Initially, several of the eroding bodies are closer to the outer wall than the $5h$ threshold \nick{required \st{to guarantee that} for} the trapezoid rule to achieve machine
precision.  In particular, the distance between bodies 1, 6, 13, and 15
and the outer wall is $1.3h$, $2.9h$, $2.8h$, and $1.3h$, respectively,
where $h$ is the arclength spacing of the outer wall $\Gamma$.  In
addition, the distance between several pairs of eroding bodies,
including 1 \& 6, 3 \& 9, 6 \& 8, and 14 \& 18, is too small to be
accurately resolved with the trapezoid rule.  By using the Barycentric
quadrature rule, the interaction between these nearly-touching bodies
can be accurately resolved \nick{\st{at the prescribed resolution} to the desired accuracy}, and erosion can
be simulated until all the bodies have vanished.

@@

\nick{\st{Similar to our previous work,}} Erosion causes the space between nearly
touching bodies to quickly expand, and flat faces develop along the
regions of near contact.  This qualitative behavior is seen in Figure~\ref{fig:Eroding20vort} between bodies 3 \& 4, 15 \& 16, and
\nick{was also witnessed in previous work}.  
However, by resolving the interaction between bodies that are
much closer together than in previous work, we observe that very little erosion occurs between certain pairs of bodies, at least initially.  
For instance the opening between bodies 1 \& 6, 3 \& 9, and 5
\& 13 grow much slower than the opening between bodies 15 \& 16.  A
common feature of the pores that grow slowly is that they are nearly
perpendicular to the main flow direction, \nick{resulting} in a small
erosion rate.

Following our previous work~\cite{qua-moo2018}, we begin by analyzing
the effect of erosion on the area fraction and the flow rate.  In
Figure~\ref{fig:Eroding20flowrate}(a), we plot the area fraction as a
function of normalized time.  The general trend of the area fraction
resembles our previous work~\cite{qua-moo2018} (Figure 10(a)), but with
a larger initial area fraction.  In
Figure~\ref{fig:Eroding20flowrate}(b), we plot the flow rate $U$
required to maintain a constant pressure drop across the channel.
Again, the trend of $U$ resembles that of our previous
work~\cite{qua-moo2018} (Figure 10(b)), except that the initial flow
rate is an order of magnitude smaller because of the larger initial area
fraction.  Starting around normalized time $0.2$,
Figure~\ref{fig:Eroding20flowrate}(b) \nick{is roughly linear which indicates
that the flow rate can be written as a power law.  Using a line of best
fit, the flow rate is approximately 
\begin{align} 
  U \approx \left(\frac{t}{t_f}\right)^{3.43},
\end{align}
which is the dashed line in} Figure~\ref{fig:Eroding20flowrate}(b).

\todo[inline]{WAIT.... Fig 7b is a semi-log plot with the x-axis linear and the y-axis logarithmic. Therefore, a straight line indicates exponential increase, not a power law.}

@@

\begin{figure}[H]
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_1}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_46}
\caption{100 bodies, 258 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_112}
\caption{97 bodies, 249 pores}
\end{subfigure}
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_164}
\caption{92 bodies, 234 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_207}
\caption{75 bodies, 189 pores}
\end{subfigure}%
\begin{subfigure}[b]{0.3\textwidth}
\includegraphics*[width =\linewidth]{./figs/hist100b_246}
\caption{34 bodies, 74 pores}
\end{subfigure}
\caption{\label{fig:Eroding100gap_hist} The distribution of the pore
sizes of 100 eroding bodies at six porosities. The black curves are
Weibull distributions whose first and second moments agree with the
data. The porosities and Weibull distribution parameters $(k,\lambda)$
at each of the porosities are: (a) $\phi = 50.09\%$ and
$(k,\lambda)=(1.4650,0.0605)$; (b) $\phi = 55.02\%$ and
$(k,\lambda)=(1.9485,0.0737)$; (c) $\phi = 65.03\%$ and
$(k,\lambda)=(2.5682,0.0965)$; (d) $\phi = 75.14\%$ and
$(k,\lambda)=(2.7771,0.1255)$; (e) $\phi = 85.07\%$ and
$(k,\lambda)=(2.6235,0.1713)$; (f) $\phi = 95.12\%$ and
$(k,\lambda)=(1.9840, 0.3194)$.}
\end{figure}
